Are Transformed Low Discrepancy Points Also Low Discrepancy?

### 2019-05-26 InvPts_verB.tex
Changes from remote

\newcommand{\blind}{1}
\newtheorem{lem}{Lemma}
\newtheorem{thm}{Theorem}
\newtheorem{pro}{Proposition}
\newtheorem{defn}{Definition}
\newtheorem{rmk}{Remark}
\newtheorem{cor}{Corollary}
\newcommand{\reals}{\mathbb{R}}
\newcommand{\naturals}{\mathbb{R}}
\newcommand{\va}{\boldsymbol{a}}
\newcommand{\vb}{\boldsymbol{b}}
\newcommand{\vc}{\boldsymbol{c}}
\newcommand{\vx}{\boldsymbol{x}}
\newcommand{\vX}{\boldsymbol{X}}
\newcommand{\vt}{\boldsymbol{t}}
\newcommand{\vz}{\boldsymbol{z}}
\newcommand{\vu}{\boldsymbol{u}}
\newcommand{\vv}{\boldsymbol{v}}
\newcommand{\vy}{\boldsymbol{y}}
\newcommand{\fraku}{\mathfrak{u}}
\newcommand{\frakv}{\mathfrak{v}}
\newcommand{\mX}{{\mathsf X}}
\newcommand{\vinfty}{\boldsymbol{\infty}}
\newcommand{\vzero}{\boldsymbol{0}}
\newcommand{\vPsi}{\boldsymbol{\Psi}}
\newcommand{\dif}{{\rm d}}
\DeclareMathOperator{\sign}{sign}
\DeclareMathOperator*{\argmax}{argmax}
\newcommand{\Udes}{\mathcal{U}}
\newcommand{\Xdes}{\mathcal{X}}
\newcommand{\FXdes}{F_{\Xdes}}
\newcommand{\cm}{\mathcal{M}}
\newcommand{\ch}{\mathcal{H}}
\newcommand{\MATLAB}{\textsc{Matlab}\xspace}
\newcommand{\Ftar}{F}
\newcommand{\ftar}{\varrho}
\newcommand{\cube}{\ensuremath{(0,1)^d}}
\newcommand{\bbone}{\mathbbm{1}}
\newcommand{\Ex}{\mathbb{E}}
\newcommand{\ip}[3][{}]{\ensuremath{\left \langle #2, #3 \right \rangle_{#1}}}
\newcommand{\norm}[2][{}]{\ensuremath{\left \lVert #2 \right \rVert}_{#1}}
\DeclareMathOperator{\cov}{cov}
\DeclareMathOperator{\err}{err}
\newcommand{\unif}{\textup{unif}}
\newcommand{\normal}{\textup{normal}}
\newcommand{\frakd}{\mathfrak{d}}
\newcommand{\tK}{\widetilde{K}}
\newcommand{\tOmega}{\widetilde{Omega}}

@@

First, we define the point deletion function, $\frakd_p$, as the change in square discrepancy resulting from removing the a point from the design:
\begin{equation}\label{eq:rowdeletion}
\frakd_p(i) = D^2(\Xdes)-D^2(\Xdes\backslash \{\vx_i\}).
\end{equation}
Here, the design $\Xdes\backslash \{\vx_i\}$ is the $N-1$ point design with the point $\{\vx_i\}$ removed. We suppress the choice of target distribution and kernel in the above discrepancy notation for simplicity. We then choose 
\begin{equation} \label{eq:rowchoice}
    i^*=\argmax_{i=1,\ldots,N} \frakd_p(i).
\end{equation}
If $\frakd_p(i^*) > 0$, it means that the smaller design with $\vx_{i^*}$ removed has a smaller discrepancy. Thus, it is reasonable to assume that moving $\vx_{i^*}$ might lead to a smaller discrepancy. On the other hand, if $\frakd_p(i^*)< 0$, it means that removing any $\vx_i$ makes the discrepancy bigger, but removing $\vx_{i^*}$ does the least damage.  Again,  $\vx_{i^*}$ is a prime candidate for modification.

Next, we define a coordinate deletion function, $\frakd_{c}$, as the change in the square discrepancy resulting from removing a coordinate in our calculation of the discrepancy:
\begin{equation}\label{eq:coorddeletion}
\frakd_c(k) = D^2(\Xdes)-D^2(\Xdes_{-k}).
\end{equation}
Here, the design $\Xdes_{-k}$ still has $N$ points but now only $d$ dimensions, the $k^{\text{th}}$ coordinate having been removed.  For this calculation to be feasible, we need the target distribution to independent marginals defined on the domain.  Also, the kernel to be of product form.  To simplify the notation, we assume a somewhat stronger condition, namely that each term in the product defining the kernel is the same for every coordinate:
\begin{equation}
    \label{eq:prodkernel}
    \Omega = \tOmega \times \cdots \times \tOmega, \qquad K(\vt,\vx) = \prod_{k=1}^d \tK(t_k,x_k), \qquad \tK:\tOmega \times \tOmega \to \reals.
\end{equation}
where the term in the product is 
We then choose 
\begin{equation} \label{eq:rowchoice}
    k^*=\argmax_{k=1, \ldots, d} \frakd_c(k).
\end{equation}
For reasons analogous to those given above, the $k^{*\text{th}}$ coordinate seems the best candidate for change.

@@

For kernels of product form, \eqref{eq:prodkernel}, and target distributions with independent marginals, the formula for the squared discrepancy in \eqref{eq:discDef} becomes
\begin{align}
\nonumber
    D^2(\Xdes,\rho,K)  & = \prod_{k=1}^d c_k  - \frac 2N \sum_{i=1}^N \prod_{k=1}^d h_k(x_{ik}) + \frac{1}{N^2}  \sum_{i,j=1}^N \prod_{k=1}^d K_k(x_{ik},x_{jk}),
    \intertext{where}
    c_k & = \int_{\Omega_k \times \Omega_k} \tK_k(t_k,x_k) \, \varrho_k(t) \varrho_k (x) \, \dif t\dif x, \\
    h_k(x) & = \int_{\Omega_k} \tK_k(t,x)  \, \varrho_k(t) \, \dif t.
\end{align}
Then the point deletion function becomes fred
\begin{align}\label{eq:rowdeletionA}
\frakd_p(i) 
&= \frac{2}{N}\biggl[ \frac{1}{N-1} \sum_{\substack{j=1 \\ j \ne i}}^N \prod_{k=1}^d h_k(x_{ik}) - \prod_{k=1}^d h_k(x_{ik})  \biggr] \\
& \qquad \qquad + \frac{1}{N^2}\biggl[\frac{1-2N}{(N-1)^2}\sum_{\substack{j,m=1 \\ j,m \ne i}}^N \tK_k(x_{jk},x_{mk}) + \sum_{j=1}^N\prod_{s=1}^d h(x_{is},x_{js})-\prod_{s=1}^d h(x_{is},x_{is})\biggr]. 
\end{align}
We use $D(\Xdes)$ as a simplified notation for the discrepancy as the target design is always standard normal distribution and the kernel function is \eqref{eq:OrigKernel}. 
The notation $\Xdes\backslash \vx_i$ means that the design $\Xdes$ without the $i$-th point $\vx_i$. 
Here $c(x)$ and $h(x,y)$ are 
\begin{align*}
c(x) &= 1+\frac{1}{\sqrt{2\pi}}+\frac{1}{2}|x|-x[\Phi_1(x)-1/2]-\phi(x),\\
h(x,y) &= 1+\frac{1}{2}|x|+\frac{1}{2}|y|-\frac{1}{2}|x-y|.
\end{align*}
The definition of $d_r(i)$ represents the change in $D^2(\Xdes)$ when $\vx_i$ is removed from the design. 
If $d_r(i)\geq 0$, it means that removing $\vx_i$ would lead to a smaller discrepancy. 
Thus, it is reasonable to assume $\vx_i$ has the potential to be updated and leads to a smaller discrepancy. 
Among all the design points that $d_r(i)\geq 0$, we should pick the $i_0$ that has largest $d_r(i)$, because removing $\vx_i$ leads to the biggest improvement, so we consider this row has the biggest potential. 
On the other hand, if $d_r(i)< 0$, it means that removing $\vx_i$ would make the discrepancy bigger, and thus make the design worse, so we should keep $\vx_i$ in the design. 
However, if $d_r(\vx_i)\leq 0$ for $i=1,\ldots, N$, and we must pick a row index, then we should pick the $i_0$ that has the smallest $\abs{d_r(\vx_i)}$ because removing $\vx_{i_0}$ would make the design worse with the least degree compared to other indices. 
Combining both scenarios, we should pick the row index by 
\[
i^*=\argmax_{i=1,\ldots,N} d_r(i).
\]
Similarly, we define the column deletion function as follows. 
\begin{eqnarray}\label{eq:coldeletion}
d_c(j) &=& -\frac{2}{N}\sum_{i=1}^N(c(x_{ij})-1)\prod\limits_{\substack{s=1,\\ s\neq j}}^dc(x_{is})\nonumber\\ 
&&+\frac{1}{N^2}\sum_{\substack{i=1,\\ r=1}}^N\left(h(x_{ij},x_{rj})-1\right)\prod_{\substack{s=1\\s\neq j}}^d h(x_{is},x_{rs}).
\end{eqnarray}
We pick the column index by 
\[
j^*=\argmax_{j=1,\ldots,d} d_c(j).
\]
Once the $(i^*, j^*)$-th coordinate is identified to be changed, we compute the delta function $\Delta(x_{i^*j^*},x)$, which represents the decrease in $D^2(\Xdes)$, if $x_{i^*j^*}$ is replaced by $x \in \mathbb{R}$. 
The delta function can be computed as follows. 
\begin{eqnarray}\label{eq:deltafunction}
&&\Delta(x_{{i^*j^*}}, x)\nonumber\\
&=& -\frac{2}{N}\left(c(x_{i^*j^*})-c(x)\right)\prod\limits_{\substack{j=1,\\j\neq j^*}}^dc(x_{i^*j})\nonumber\\
&&+\frac{1}{N^2}\left(2\sum_{\substack{i=1,\\i\neq i^*}}^N(h(x_{i^*j^*},x_{ij^*})-h(x,x_{ij^*}))\prod_{\substack{j=1,\\j\neq j^*}}^dh(x_{i^*j},x_{ij})\right.\nonumber\\
&&\left.+(|x_{i^*j^*}|-|x|)\prod_{\substack{j=1,\\j\neq j^*}}^d[1+|x_{i^*j}|]\right)
\end{eqnarray} 
We summarize the coordinate-exchange algorithm in Algorithm 1. 
\begin{algorithm}[ht]
\begin{algorithmic}[1]%\label{alg:coorex}
\caption{Coordinate exchange method}
\State Set $k=0$.
\State Compute the row deletion function $d_r(i)$ ($i=1,...,N$) for $N$ rows of $\mX^{(k)}$ and choose the row $i^*$ with the largest row deletion value, i,e, $i^* = \argmax_{i} d_r(i)$. 
\State Compute the column deletion function $d_c(j)$ ($j=1,...,d$) for $d$ columns of $\mX^{k}$ and choose the column $j^*$ with the largest column deletion value, i.e., $j^* = \argmax_{j} d_c(j)$. 
\State Solve the univariate nonlinear optimization problem
$$\max_{x} \Delta(x_{i^*j^*},x),\,\,\,\,\text{s.t.} x\in \mathbb{R},$$
and denote the optimal solution as $x^*$.
\If{$\Delta(x_{i^*j^*},x^*)>0$} 
\State {replace $x_{i^*j^*}$ with $x^*$ and the define the new design as $\mX^{(k+1)}$.}
\State {$k=k+1$.}
\Else 
\State{terminate the algorithm and conclude the current design is optimal.} 
\EndIf
\end{algorithmic}
\end{algorithm}

>>>>

\newcommand{\blind}{1}
\newtheorem{lem}{Lemma}
\newtheorem{thm}{Theorem}
\newtheorem{pro}{Proposition}
\newtheorem{defn}{Definition}
\newtheorem{rmk}{Remark}
\newtheorem{cor}{Corollary}
\newcommand{\reals}{\mathbb{R}}
\newcommand{\naturals}{\mathbb{R}}
\newcommand{\va}{\boldsymbol{a}}
\newcommand{\vb}{\boldsymbol{b}}
\newcommand{\vc}{\boldsymbol{c}}
\newcommand{\vx}{\boldsymbol{x}}
\newcommand{\vX}{\boldsymbol{X}}
\newcommand{\vt}{\boldsymbol{t}}
\newcommand{\vz}{\boldsymbol{z}}
\newcommand{\vu}{\boldsymbol{u}}
\newcommand{\vv}{\boldsymbol{v}}
\newcommand{\vy}{\boldsymbol{y}}
\newcommand{\fraku}{\mathfrak{u}}
\newcommand{\frakv}{\mathfrak{v}}
\newcommand{\mX}{{\mathsf X}}
\newcommand{\vinfty}{\boldsymbol{\infty}}
\newcommand{\vzero}{\boldsymbol{0}}
\newcommand{\vPsi}{\boldsymbol{\Psi}}
\newcommand{\dif}{{\rm d}}
\DeclareMathOperator{\sign}{sign}
\DeclareMathOperator*{\argmax}{argmax}
\newcommand{\Udes}{\mathcal{U}}
\newcommand{\Xdes}{\mathcal{X}}
\newcommand{\FXdes}{F_{\Xdes}}
\newcommand{\cm}{\mathcal{M}}
\newcommand{\ch}{\mathcal{H}}
\newcommand{\MATLAB}{\textsc{Matlab}\xspace}
\newcommand{\Ftar}{F}
\newcommand{\ftar}{\varrho}
\newcommand{\cube}{\ensuremath{(0,1)^d}}
\newcommand{\bbone}{\mathbbm{1}}
\newcommand{\Ex}{\mathbb{E}}
\newcommand{\ip}[3][{}]{\ensuremath{\left \langle #2, #3 \right \rangle_{#1}}}
\newcommand{\norm}[2][{}]{\ensuremath{\left \lVert #2 \right \rVert}_{#1}}
\DeclareMathOperator{\cov}{cov}
\DeclareMathOperator{\err}{err}
\newcommand{\unif}{\textup{unif}}
\newcommand{\normal}{\textup{normal}}
\newcommand{\frakd}{\mathfrak{d}}
\newcommand{\tK}{\widetilde{K}}
\newcommand{\tOmega}{\widetilde{\Omega}}
\newcommand{\Order}{\mathcal{O}}

@@

First, we define the point deletion function, $\frakd_p$, as the change in square discrepancy resulting from removing the a point from the design:
\begin{equation}\label{eq:rowdeletion}
\frakd_p(i) = D^2(\Xdes)-\left(\frac{N-1}{N}\right)^2 D^2(\Xdes\backslash \{\vx_i\}).
\end{equation}
Here, the design $\Xdes\backslash \{\vx_i\}$ is the $N-1$ point design with the point $\{\vx_i\}$ removed. We suppress the choice of target distribution and kernel in the above discrepancy notation for simplicity. We then choose 
\begin{equation} \label{eq:rowchoice}
    i^*=\argmax_{i=1,\ldots,N} \frakd_p(i).
\end{equation}
The definition of $i^*$ means that removing $\vx_{i^*}$ from the design $\Xdes$ results in the smallest discrepancy among all possible deletions.  Thus, $\vx_{i^*}$ is helping the least, which makes it a prime candidate for modification.

Next, we define a coordinate deletion function, $\frakd_{c}$, as the change in the square discrepancy resulting from removing a coordinate in our calculation of the discrepancy:
\begin{equation}\label{eq:coorddeletion}
\frakd_c(k) = D^2(\Xdes)-D^2(\Xdes_{-k}).
\end{equation}
Here, the design $\Xdes_{-k}$ still has $N$ points but now only $d$ dimensions, the $k^{\text{th}}$ coordinate having been removed.  For this calculation to be feasible, we need the target distribution to independent marginals defined on the domain.  Also, the kernel to be of product form.  To simplify the notation, we assume a somewhat stronger condition, namely that each term in the product defining the kernel is the same for every coordinate:
\begin{equation}
    \label{eq:prodkernel}
    \Omega = \tOmega \times \cdots \times \tOmega, \qquad K(\vt,\vx) = \prod_{k=1}^d [1+ \tK(t_k,x_k)], \qquad \tK:\tOmega \times \tOmega \to \reals.
\end{equation}
where the term in the product is 
We then choose 
\begin{equation} \label{eq:rowchoice}
    k^*=\argmax_{k=1, \ldots, d} \frakd_c(k).
\end{equation}
For reasons analogous to those given above, the $k^{*\text{th}}$ coordinate seems the best candidate for change.

@@

For kernels of product form, \eqref{eq:prodkernel}, and target distributions with independent marginals, the formula for the squared discrepancy in \eqref{eq:discDef} becomes
\begin{align}
\nonumber
    D^2(\Xdes,\rho,K)  & = c^d  - \frac 2N \sum_{i=1}^N H(\vx_{i}) + \frac{1}{N^2}  \sum_{i,j=1}^N K(\vx_{i},x_{j}),
    \intertext{where}
    c & = \int_{\Omega_k \times \Omega_k} \tK_k(t_k,x_k) \, \varrho_k(t) \varrho_k (x) \, \dif t\dif x, \\
    H(\vx) = \prod_{k=1}^d [1+ h(x_{k})]
    h(x) & = \int_{\Omega_k} \tK_k(t,x)  \, \varrho_k(t) \, \dif t.
\end{align}
For a standard multivariate normal target distribution and the kernel defined in \eqref{eq:OrigKernel}, we have
\begin{align*}
c &= 1+\sqrt{\frac{2}{\pi}}, \\
h(x) &= \frac{1}{\sqrt{2\pi}}+\frac{1}{2}|x|-x[\Phi(x)-1/2]-\phi(x),\\
\tK(t,x) &= \frac{1}{2} (|t|+ |x|- |x-t|).
\end{align*}

The point deletion function defined in \eqref{eq:rowdeletion} then can be expressed as
\begin{align}
\nonumber
\frakd_p(i) 
&= \frac{(2N-1)c^d}{N^2} - \frac{2}{N}\biggl[ \frac{1}{N} \sum_{j=1}^N  H(\vx_{j}) + \left(1 - \frac 1N \right) H(\vx_{i})  \biggr] \\
\label{eq:rowdeletionA}
& \qquad \qquad + \frac{1}{N^2}\biggl[2 \sum_{j=1}^N K(\vx_{i},\vx_{j})-\prod_{k=1}^d K(\vx_{i},\vx_{i})\biggr]. 
\end{align}
The computational cost for $\frakd_p(1), \ldots, \frakd_p(N)$ is then $\Order(dN^2)$, which is the same order as the cost of the discrepancy of a single design.

We use $D(\Xdes)$ as a simplified notation for the discrepancy as the target design is always standard normal distribution and the kernel function is \eqref{eq:OrigKernel}. 
The notation $\Xdes\backslash \vx_i$ means that the design $\Xdes$ without the $i$-th point $\vx_i$. 
Here $c(x)$ and $h(x,y)$ are 
\begin{align*}
c(x) &= 1+\frac{1}{\sqrt{2\pi}}+\frac{1}{2}|x|-x[\Phi_1(x)-1/2]-\phi(x),\\
h(x,y) &= 1+\frac{1}{2}|x|+\frac{1}{2}|y|-\frac{1}{2}|x-y|.
\end{align*}
The definition of $d_r(i)$ represents the change in $D^2(\Xdes)$ when $\vx_i$ is removed from the design. 
If $d_r(i)\geq 0$, it means that removing $\vx_i$ would lead to a smaller discrepancy. 
Thus, it is reasonable to assume $\vx_i$ has the potential to be updated and leads to a smaller discrepancy. 
Among all the design points that $d_r(i)\geq 0$, we should pick the $i_0$ that has largest $d_r(i)$, because removing $\vx_i$ leads to the biggest improvement, so we consider this row has the biggest potential. 
On the other hand, if $d_r(i)< 0$, it means that removing $\vx_i$ would make the discrepancy bigger, and thus make the design worse, so we should keep $\vx_i$ in the design. 
However, if $d_r(\vx_i)\leq 0$ for $i=1,\ldots, N$, and we must pick a row index, then we should pick the $i_0$ that has the smallest $\abs{d_r(\vx_i)}$ because removing $\vx_{i_0}$ would make the design worse with the least degree compared to other indices. 
Combining both scenarios, we should pick the row index by 
\[
i^*=\argmax_{i=1,\ldots,N} d_r(i).
\]
Similarly, we define the column deletion function as follows. 
\begin{eqnarray}\label{eq:coldeletion}
d_c(j) &=& -\frac{2}{N}\sum_{i=1}^N(c(x_{ij})-1)\prod\limits_{\substack{s=1,\\ s\neq j}}^dc(x_{is})\nonumber\\ 
&&+\frac{1}{N^2}\sum_{\substack{i=1,\\ r=1}}^N\left(h(x_{ij},x_{rj})-1\right)\prod_{\substack{s=1\\s\neq j}}^d h(x_{is},x_{rs}).
\end{eqnarray}
We pick the column index by 
\[
j^*=\argmax_{j=1,\ldots,d} d_c(j).
\]
Once the $(i^*, j^*)$-th coordinate is identified to be changed, we compute the delta function $\Delta(x_{i^*j^*},x)$, which represents the decrease in $D^2(\Xdes)$, if $x_{i^*j^*}$ is replaced by $x \in \mathbb{R}$. 
The delta function can be computed as follows. 
\begin{eqnarray}\label{eq:deltafunction}
&&\Delta(x_{{i^*j^*}}, x)\nonumber\\
&=& -\frac{2}{N}\left(c(x_{i^*j^*})-c(x)\right)\prod\limits_{\substack{j=1,\\j\neq j^*}}^dc(x_{i^*j})\nonumber\\
&&+\frac{1}{N^2}\left(2\sum_{\substack{i=1,\\i\neq i^*}}^N(h(x_{i^*j^*},x_{ij^*})-h(x,x_{ij^*}))\prod_{\substack{j=1,\\j\neq j^*}}^dh(x_{i^*j},x_{ij})\right.\nonumber\\
&&\left.+(|x_{i^*j^*}|-|x|)\prod_{\substack{j=1,\\j\neq j^*}}^d[1+|x_{i^*j}|]\right)
\end{eqnarray} 
We summarize the coordinate-exchange algorithm in Algorithm 1. 
\begin{algorithm}[ht]
\begin{algorithmic}[1]%\label{alg:coorex}
\caption{Coordinate exchange method}
\State Set $k=0$.
\State Compute the row deletion function $d_r(i)$ ($i=1,...,N$) for $N$ rows of $\mX^{(k)}$ and choose the row $i^*$ with the largest row deletion value, i,e, $i^* = \argmax_{i} d_r(i)$. 
\State Compute the column deletion function $d_c(j)$ ($j=1,...,d$) for $d$ columns of $\mX^{k}$ and choose the column $j^*$ with the largest column deletion value, i.e., $j^* = \argmax_{j} d_c(j)$. 
\State Solve the univariate nonlinear optimization problem
$$\max_{x} \Delta(x_{i^*j^*},x),\,\,\,\,\text{s.t.} x\in \mathbb{R},$$
and denote the optimal solution as $x^*$.
\If{$\Delta(x_{i^*j^*},x^*)>0$} 
\State {replace $x_{i^*j^*}$ with $x^*$ and the define the new design as $\mX^{(k+1)}$.}
\State {$k=k+1$.}
\Else 
\State{terminate the algorithm and conclude the current design is optimal.} 
\EndIf
\end{algorithmic}
\end{algorithm}
